On the presentation of orthogonal groups over polynomial rings

### 2019-01-29 dl-paper.tex
More stuff added to Proposition 1

\subsection{Unstable Quillen K-groups}
\begin{comment}
In order to better understand $C(\Phi, R, I)$  we would like to continue the sequence~\eqref{eq:relative-St} to the left. 
Thus we need to work with unstable $\K_3$-functors a la Quillen (cf.~\cite{Pa89}).
\end{comment}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following assumptions:
\begin{enumerate}
 \item\label{req:left-exact} $G$ is left exact;
 \item $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} $S$ preserves coqualizers of kernel pairs.
\end{enumerate}

@@

\begin{prop}\label{characterization}
 Then the following holds:
\begin{proplist}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item There is an exact sequence natural in $(R, I)$:
 \end{proplist}
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & \HH_3(S(R), S(R/I), \ZZ).} \]
\end{prop}
\begin{proof}
 The first claim follows from the solution of Exercises~1.8--1.9 of \cite[Chapter~IV]{Kbook}. %provided one verifies that all the intermediate isomorphisms are natural in $R$.

 Let us verify the second claim.
 Denote by $p_0, p_1, \Delta$ the two obvious projections and the diagonal map $\xymatrix{D_{R,I} \ar@<1pt>[r] \ar@<-2.5pt>[r] & \ar@<-4pt>[l] R}$
  and set $d_i = Sp_i$, $s = S\Delta$.
 Obviously, $d_0s = d_1 s = id_{S(R)}$. Set $G_i = \Ker(d_i)$, $S(R, I) = G_0 / [G_0, G_1]$ and denote by $\mu$ the natural map $S(R, I) \to N$ induced by $d_1$.

 From~\eqref{req:left-exact} and $\varphi(G_i) \subseteq \Ker(G(p_i))$ we conclude that $G_0 \cap G_1 \subseteq \Ker(\varphi) \colon S(D_{R, I}) \to G(D_{R, I})$
  and hence $G_0 \cap G_1$ is a central subgroup of $S(D_{R, I})$.
 Now invoking~\cite[Proposition~6]{Lo78} we conclude that the extension $S(R, I) \to S(R)$ is a universal relative central extension of the coequalizer $\mathrm{coeq}(d_0, d_1)$, 
   which coincides with $S(R) \to S(R/I)$ by~\eqref{req:coeq}.

 By~\cite[Th{\'e}or{\`e}me~2]{Lo78} the relative homology group $\HH_3(\St(\Phi, R/I), \St(\Phi, R), \ZZ)$ can be identified
 with the kernel $C(\Phi, R, I)$ of the relative central extension $\St(\Phi, R, I) \to \St(\Phi, R)$.
\end{proof}

>>>>

\subsection{Unstable Quillen K-groups}
\begin{comment}
In order to better understand $C(\Phi, R, I)$  we would like to continue the sequence~\eqref{eq:relative-St} to the left. 
Thus we need to work with unstable $\K_3$-functors a la Quillen (cf.~\cite{Pa89}).
\end{comment}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following assumptions:
\begin{enumerate}
 \item\label{req:left-exact} $G$ is left exact;
 \item $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} $S$ preserves coequalizers of kernel pairs.
\end{enumerate}

@@

Let us also briefly recall Loday's procedure for relativization of $S$, that is,
 extending $S$ from the category of rings to the category of pairs.
Denote by $p_0, p_1, \Delta$ the two obvious projections and the diagonal map $\xymatrix{D_{R,I} \ar@<1pt>[r] \ar@<-2.5pt>[r] & \ar@<-4pt>[l] R}$
  and set $d_i = Sp_i$, $s = S\Delta$.
 Obviously, $d_0s = d_1 s = id_{S(R)}$. Set $G_i = \Ker(d_i)$, and define $S(R, I)$ as $ G_0 / [G_0, G_1]$, denote by $\mu$ the natural map $S(R, I) \to S(R)$ induced by $d_1$.

\begin{prop}\label{characterization}
\begin{proplist}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item The map $\mu \colon S(R, I) \to S(R)$ is a universal relative central extension of 
    $S(R) \to S(R/I)$, whose kernel is naturally isomorphic to the third relative homology group $\HH_3(S(R), S(R/I), \ZZ)$.
 \item There is an exact sequence natural in $(R, I)$:
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & \Ker(\mu).} \]
 \end{proplist}   
\end{prop}
\begin{proof}
 The first claim follows from the solution of Exercises~1.8--1.9 of \cite[Chapter~IV]{Kbook}. %provided one verifies that all the intermediate isomorphisms are natural in $R$.

 Let us verify the second claim. From~\eqref{req:left-exact} and $\varphi(G_i) \subseteq \Ker(G(p_i))$ we conclude that $G_0 \cap G_1 \subseteq \Ker(\varphi \colon S(D_{R, I}) \to G(D_{R, I}))$
  and hence $G_0 \cap G_1$ is a central subgroup of $S(D_{R, I})$.
 Now invoking~\cite[Proposition~6]{Lo78} we conclude that the extension $S(R, I) \to S(R)$ is a universal relative central extension of the coequalizer $\mathrm{coeq}(d_0, d_1)$, 
   which coincides with $S(R) \to S(R/I)$ by~\eqref{req:coeq}. The kernel of $\mu$ is isomorphic to $\HH_3(S(R), S(R/I), \ZZ)$ by~\cite[Th{\'e}or{\`e}me~2]{Lo78}.

 The third claim now follows from the starting portion of the homology long exact sequence of the pair $(S(R/I), S(R))$ and the first two assertions.
\end{proof}
